---
title: "Is polynomial regression nonlinear?"
format:
  html:
    theme: default
---
Load libraries & generate data.  

```{r}
pacman::p_load(dplyr, ggplot2, scatterplot3d)


set.seed(123)
x = runif(100)
y = runif(100)
x_sq = x^2

plot(x, x_sq)
```

Regress $y$ on $x$ and $x^2$.  
Save predictions and all variables in a data frame. 

```{r}
df= data.frame(x, x_sq, y)
lm1 = lm(y ~ x + x_sq, data = df)
#add predictions to df
df = data.frame(y_pred = predict(lm1, df), df)
  (lm1)
```

Plot $y$ and $x$, and model relationship as nonlinear based on the linear model.
```{r}
df %>%
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(x, y_pred)) 
#  geom_smooth(method = 'lm', formula = y ~ x + I(x^2))
```


Create plot with all three dimensions.
```{r}
s = scatterplot3d(x=df$x, y=df$x_sq, z=df$y, color="gray", pch=1, 
                  xlab="X", ylab="X^2", zlab="Y", 
                  xlim=c(0, 1), ylim=c(0,1), 
                  zlim=c(0, 1), type="h", main="In pseudo-3D space")
s$points(x=df$x, y=df$x_sq, z=df$y, col="red", pch=1)
# add in paramters from lm1 model:
s$plane3d(Intercept=0.6974, 
          x.coef= -0.9127, y.coef=0.8263, 
          col="blue")
```

**RESULT: a linear (flat) plane**
- So even though we are using a 'nonlinear' relationship to model X and Y, we are still technically using a linear model.
- The hyperplane that exists in the 3 dimensional space is flat/linear.
- The 2d version we see on the typical graph, used to think about the relationship between x and y, appears nonlinear.




